---
title: "Transform"
execute:
  echo: true
  eval: false
  message: false
  warning: false
  freeze: auto
---

## Dates

### Format date
```{r}
as.Date(paste0("15_", "Jan_1998"), format = "%d_%b_%Y")
```

## Strings

### Extract string between strings

```{r}
library(stringr)
x <- "here and now"
str_extract(x, "(?<=here).*(?=now)")
```

## Missing values

### Count number of missing values in each column
```{r}
df |> map_dbl(~sum(is.na(.))) colSums(is.na(x = df)))
```

### Count unique values in every column
```{r}
df |> map_int(n_distinct)
```

### Show unique values in categorical variables
```{r}
map(select_if(df, is.character), unique)
```

### Convert selected string in single column to NA
```{r}
df |> mutate(col1 = na_if(col1, "Unknown"))
df |> mutate_at(vars(col1, col2), na_if, "Unknown")
```

### Convert multiple strings across all columns to NA
```{r}
df |> mutate(across(where(is.character), ~replace(., . %in% c("Unknown","unknown"), NA)))
```

### Convert string to NA across all columns
```{r}
mutate(across(everything(), na_if, ".."))
```

## Other
### Mutate and move new column
```{r}
penguins |> mutate(id = row_number(), .before = contains("_"))
```

### Standardise values
```{r}
df |> mutate(col1 = scale(col1))
```

### Round percentages
```{r}
scales::label_percent(accuracy = 1L)
```